---
title: "RasterAnalysis"
author: "Jos Goris, Mark ten Vregelaar"
date: "January 14, 2016"
output: html_document
---

Import libraries and download the data
```{r, message=FALSE}
library(raster)
library(rasterVis)

ifolder <- './data/'
ofolder <- './output/'
dir.create(ifolder, showWarnings = FALSE)
dir.create(ofolder, showWarnings = FALSE)
dataURL <- "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/archive/gh-pages.zip"
inputZip <- list.files(path=ifolder, pattern= '^.*\\.zip$')

if (length(inputZip) == 0){ ##only download when not alrady downloaded
  download.file(url = dataURL, destfile = 'data/landsatData.zip', method = 'wget')
} 

unzip('data/landsatData.zip', exdir=ifolder)
```

Load the data
```{r}
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB2.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB3.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB4.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB1.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB5.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB7.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/vcfGewata.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/trainingPoly.rda")
```

Brick the layers
```{r}
gewata <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7)
```

Remove all values above 100 for 
```{r}
vcfGewata[vcfGewata > 100] <- NA
```

Add VCF to brick
```{r}
covs <- addLayer(gewata, vcfGewata)
```

Rename the columns
```{r}
names(covs) <- c('band1', 'band2', 'band3', 'band4', 'band5', 'band7', 'vcf')
```

Pair the bands to the vcf
```{r}
pairs(covs)
```

Make covs a data frame and omit the NA's
```{r}
valuetable <- getValues(covs) 
valuetable <- na.omit(valuetable)
valuetable <- as.data.frame(valuetable)
```

Make the linear model and show the summary
```{r}
regression <- lm(vcf ~ band1 + band2 + band3 + band4 + band5 + band7, data = valuetable)
summary(regression)
```

Predict the VCF
```{r}
predVCF <- predict(covs, model=regression, na.rm=TRUE)
```

Remove the negative values
```{r}
predVCF[predVCF<0] <- NA
```

Plot the results
```{r}
mycolorkey <- list(at=seq(0, 100, 5), labels=list(labels=seq(0, 100, 10), at=seq(0, 100, 10)), space = 'bottom')
levelplot(stack(vcfGewata, predVCF), col.regions = colorRampPalette(c('brown','yellow','darkgreen'))(255), colorkey = mycolorkey,
          names.attr = c('Original VCF','Predicted VCF'), main = list('Tree cover in Gewata',cex=1.8,vjust=-1),scales=list(draw=FALSE),xlab=list('Tree cover (%)',vjust=7))

```

Calculate the difference between the tree cover and the predicted tree cover
```{r}
difference <- (predVCF - vcfGewata)
```

Calculate the root mean square error
```{r}
RMSE <- sqrt((cellStats((difference)^2, stat = mean, na.rm = TRUE)))
names(RMSE) <- 'RMSE Total Raster'
```

Add column 'Code' to the trainingPoly 
```{r}
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
```

Rasterize the trainingPoly
```{r}
zones <- rasterize(trainingPoly, difference, field='Code')
```

Mask the classes
```{r}
differenceMasked <- mask(difference, zones)
```

Compute the mean for the difference^2 for the zones
```{r}
zones <- zonal((differenceMasked)^2, zones)
```

Calculate the RSME for the zones
```{r}
RMSEzones <- sqrt(zones)
```

Plot the result, remove the column 'zone', name the rows
```{r}
rownames(RMSEzones) <- c('Croplands','Forest','Wetlands')
RMSEzones <- RMSEzones[,-1]
barplot(c(RMSEzones, RMSE), main = 'RMSE per zone', ylim = c(0, 12), col = c('lightgreen','darkgreen','lightblue','pink') )
```